Shell Model for Reconstruction and Real-Time Simulation of Thin Anatomical Structures

### 2010-03-11 MICCAI2010/paper.tex
Last update (added conclusion + few changes in intro and images)
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@103 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\title{Shell Model for Reconstruction and Simulation of Thin Anatomical Structures}
%\author{Olivier Comas, St\'ephane Cotin and Christian Duriez}
%\institute{INRIA, Alcove team, Lille, France}

@@

Shell theory allows the modelling of structure deformations when the thickness is small compared to its other dimensions~\cite{Liu03}. The key idea is to model the physical shell as if it were a surface of zero thickness, but endowed with mechanical properties in the form of elastic resistance to stretching and bending forces. \OC{Pas de shearing il me semble. Et il y a bien une thickness dans la formulation, ta phrase laisse sous-entendre qu'on la prend pas en compte.}
Rather than resorting to shell theory, previous work on medical simulation often relies on linear or angular mass-spring models as in \cite{Nedel98,Hammer08}. Yet, such models are limited in their ability to describe certain behaviour, as they do not rely on continuum mechanics: it is difficult to derive spring stiffness (in particular for angular springs) from elastic properties (Young's modulus and Poisson's ratio).
The use of fast shell based model for computer graphics and interactive simulation was recently introduced with the work of Choi et al.~\cite{Choi07}. The model relies on simplified energy functions and precomputed modal analysis. It allows fast and visually realistic results. We propose to rely an a similar approach but with more accuracy in order to be applicable to medical simulation. Our model is not based on modal analysis but uses a co-rotational formulation and polynomial shape functions. In~\cite{Comas2010ISBMS}, we present and apply this shell model to a flat deformable implant. \OC{ptet un peu trop specifique pour preserver l'anonymat non ?}

@@

Our contribution is to unify these two approaches (shell based modelling and geometric surface reconstruction) separately used in mechanical engineering and computer vision in a framework for realistic simulation of thin organ deformations. In addition, we demonstrate the ability of the technique to simulate thin anatomical membrane which enfolds a solid organ without any requirement on mesh correspondence. \OC{j'enleverais bien cette derniere phrase moi, on en fait pas la demonstration au final, on evoque juste que ca serait possible} Our method, presented in section~\ref{sec:model}, is based on shell theory and computes the internal forces using Finite Element Modelling (FEM) on curved triangles. The design of the elements relies on a polynomial interpolation function that is also used to capture and process the interactions with other models, including solid models. In section~\ref{sec:mesh} we also present an automatic process to obtain meshes from image based reconstruction that relies on the exact same polynomial function.
% in an attempt to optimise the computation using fewer elements. 
Finally the benefits of our approach (meshing of a curved surface, fast computation and possible interactions with solid models) are illustrated using various examples presented in section~\ref{sec:results}.

@@

\begin{figure}
\centering
\includegraphics[height=3cm]{images/resultsLiver}
\caption {Left: the targeteted high resolution liver mesh (8,000 triangles). Right: the one-sided Hausdorff distance error map after applying only one iteration of our algorithm to the coarse mesh (1,200 shells)}. 
\label{fig-liver}
\end{figure}

\begin{figure}
\centering
\includegraphics[height=3cm]{images/resultsAnevrism}
\caption {Left: the targeteted high resolution anevrism mesh (28,368 triangles). Right: the one-sided Hausdorff distance error map after applying only one iteration of our algorithm to the coarse mesh (1,160 shells)}.
\label{fig-anevrism}
\end{figure}

@@

Anevrism: 7.5 FPS with a timestep of 0.02s so we would need to be about 7 times faster to be real-time (easily doable on GPU)

@@

\section{Conclusion}
\section*{Acknowledgments}

>>>>

\title{Shell Model for Reconstruction and Real-Time Simulation of Thin Anatomical Structures}
%\author{Olivier Comas, St\'ephane Cotin and Christian Duriez}
%\institute{INRIA, Alcove team, Lille, France}

@@

Shell theory allows the modelling of structure deformations when the thickness is small compared to its other dimensions~\cite{Liu03}. The key idea is to model the physical shell as if it were a surface of zero thickness, but endowed with mechanical properties in the form of elastic resistance to stretching and bending forces. \OC{TODO: reformuler cette phrase pour enlever zero thickness}
Rather than resorting to shell theory, previous work on medical simulation often relies on linear or angular mass-spring models as in \cite{Nedel98,Hammer08}. Yet, such models are limited in their ability to describe certain behaviour, as they do not rely on continuum mechanics: it is difficult to derive spring stiffness (in particular for angular springs) from elastic properties (Young's modulus and Poisson's ratio).
The use of fast shell based model for computer graphics and interactive simulation was recently introduced with the work of Choi et al.~\cite{Choi07}. The model relies on simplified energy functions and precomputed modal analysis. It allows fast and visually realistic results. We propose to rely an a similar approach but with more accuracy in order to be applicable to medical simulation. Our model is not based on modal analysis but uses a co-rotational formulation and polynomial shape functions presented in~\cite{Comas2010ISBMS}.

@@

Our contribution is to unify these two approaches (shell based modelling and geometric surface reconstruction) separately used in mechanical engineering and computer vision in a framework for realistic simulation of thin organ deformations. Our method, presented in section~\ref{sec:model}, is based on shell theory and computes the internal forces using Finite Element Modelling (FEM) on curved triangles. The design of the elements relies on a polynomial interpolation function that is also used to capture and process the interactions with other models, including solid models. In section~\ref{sec:mesh} we also present an automatic process to obtain meshes from image based reconstruction that relies on the exact same polynomial function.
% in an attempt to optimise the computation using fewer elements. 
Finally the benefits of our approach (meshing of a curved surface, fast computation and possible interactions with solid models) are illustrated using various examples presented in section~\ref{sec:results}.

@@

\begin{figure}
\centering
\includegraphics[height=3cm]{images/resultsLiver}
\caption {Left: the targeteted high resolution Glisson's capsule mesh (8,000 triangles). Right: the one-sided Hausdorff distance error map after applying only one iteration of our algorithm to the coarse mesh (1,200 shells)}. 
\label{fig-liver}
\end{figure}

\begin{figure}
\centering
\includegraphics[height=3cm]{images/resultsAneurysm}
\caption {(a) the targeteted high resolution aneurysm mesh (28,368 triangles). (b): the one-sided Hausdorff distance error map on a mesh of 1,160 shells generated with our method}.
\label{fig-aneurysm}
\end{figure}

@@

@@

\section{Conclusion}
We proposed a complete framework for real-time modelling of thin anatomical structures. The novelty of our method relies on the combination of a shell finite element formulation and a geometric surface reconstruction both based on the same polynomial interpolation function used to describe the surface of shells. We also showed how contacts and interactions with the curved surfaces of shells could be handled using the same function. The efficiency of the whole approach was demonstrated through shell reconstruction and real-time modelling of a blood vessel showing signs of an aneurysm. Finally these results are very encouraging towards fine modelling of more complex situations like interactions existing between the liver and the Glisson's capsule or the simulation of an angioplasty procedure.

%\section*{Acknowledgments}
